Xolotl: An Intuitive and Approachable Neuron \& Network Simulator in MATLAB

### 2018-06-01 PaperLatex/paper.tex
General clarifications

	\texttt{xolotl} is a free and open-source neuronal simulator written in \texttt{C++} with \texttt{MATLAB} wrappers. Biophysically-detailed models of networks can be designed efficiently using an intuitive language tightly coupled to the object-based architecture of the underlying \texttt{C++} code. Models can be specified by adding conductances to compartment objects. The structure is modular, serialized, and searchable, permitting high-level programmatic control over nearly all features of the models. \texttt{C++} templates are provided for developing new conductances, compartments, and integration schemata. It also includes a customizable graphical user interface (GUI) for rapid prototyping and hand-tuning conductances in real-time. The modular structure and accessibility to all parameters, variables, and dynamics of the model network in \texttt{MATLAB} facilitate rapid construction and assessment of model networks. \texttt{xolotl} is freely available at \url{https://github.com/marderlab/xolotl}. This tool provides straightforward implementation and fast simulation of neuronal models while permitting full control over every aspect of the network and integration.

@@

\texttt{xolotl} (\url{https://github.com/marderlab/xolotl}) is a fast single-compartment and multi-compartment simulator in \texttt{C++} with \texttt{MATLAB} wrappers. Written with an emphasis on ease-of-use, \texttt{xolotl} can simulate single-compartment conductance-based models, networks of these, and detailed multi-compartment models. \texttt{xolotl} exploits a novel automatic type system, \texttt{cpplab}, which binds \texttt{MATLAB} code to \texttt{C++} header files, creating objects and classes \textit{ad libitum} in \texttt{MATLAB} which reflect the underlying object-oriented code. \texttt{xolotl} implements \texttt{cpplab} to represent the nested structure of conductance-based models, and exploits the computational efficiency of the low-level programming language to quickly integrate models. For this reason, models can be implemented entirely in \texttt{MATLAB} with a few lines of code.

Models are specified in \texttt{MATLAB} by a nested structure. The \texttt{xolotl} object contains compartments which themselves contain conductances. Synapses belong to the \texttt{xolotl} object and connect compartments together. The high-level specification supports arbitrarily large network and multi-compartment morphologies.

The software has been implemented in \texttt{MATLAB} due to its ease-of-use and popularity among neuroscientists. \texttt{cpplab} provides a powerful backend for specifying and integrating models without relying on the significantly slower and limiting \texttt{MATLAB codegen}. While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.   Minimal experience with \texttt{MATLAB} is required to use \texttt{xolotl}, and all equations and integration methods are provided transparently to the end user. No string parsing of equations is required \autocite{sherfeyDynaSimMATLABToolbox2018, stimbergBrianSecondComing2013, stimbergEquationorientedSpecificationNeural2014}.

\texttt{xolotl} comes packaged with visualization functions and a graphical user interface (GUI) for real-time manipulation of model parameters. Plotting of voltage, intracellular calcium, conductance gating functions, and time constants is provided by built-in \texttt{xolotl} methods. The GUI permits real-time tuning of any network parameters using numerical sliders in a graphical interface which displays the resultant membrane potential and intracellular calcium traces. The ease-of-use of these tools lends them to pedagogical applications and rapid exploration of toy models. This tool aims to simplify the investigation of dynamics of complex neural network models, facilitate collaborative modeling, and complement other tools being developed in the neuroinformatics community.

@@

@@

Models are specified by adding compartments and synapses to the \texttt{xolotl} object. Conductances are added to compartments and controllers can be added to conductances. This modular structure recapitulates the biophysics of the Hodgkin-Huxley formalism and obviates the need to explicitly write out equations, which in \texttt{xolotl} are contained within the conductance header files.

@@

\texttt{xolotl} relies on \texttt{cpplab} constructions, which allow the user to exploit the efficiency of low-level \texttt{C++} code. \texttt{MATLAB} treats \texttt{cpplab} objects as fully-typed variables allowing for symbolic manipulation using only the high-level programming language and graphical interfaces. \texttt{xolotl} is fast because all time-intensive code is written in \texttt{C++}. While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming. \texttt{C++} provides speed improvements beyond the benefits of translating \texttt{MATLAB} features into low-level code. For this reason, \texttt{cpplab} has been designed to provide an interface for constructing, transpiling, and compiling \texttt{C++} code to be called from within \texttt{MATLAB}. \texttt{xolotl} simulations are run entirely from \texttt{C++} executables.

\paragraph{Automatic and efficient compiling.}
\texttt{xolotl} automatically uses the MD5 algorithm to hash the network and compile a new binary and \texttt{MEX} bridge file only if needed. \texttt{MATLAB} provides a high level programmatic and graphical interface for implementing, manipulating, and visualizing models without sacrificing the enhancements of the underlying \texttt{C++} code.

@@

\paragraph{Adding features.}
The \texttt{add} function will construct a \texttt{cpplab} object and affix it to as a field in the \texttt{xolotl} structure. All compartments, conductances, synapses, and controllers are \texttt{cpplab}. Compartments add to the \texttt{xolotl} object and conductances add to compartments. Specific properties can be specified using key-value pair arguments (e.g. Figure \ref{fig:figurehh}A).

\paragraph{Finding features.}
\texttt{cpplab} comes with several features which simplify the handling of complexly-nested models. The \texttt{find} function acquires a cell array of all properties of the network which satisfy a search condition. For example, one can find all paths to maximal conductances within the "'HH'" compartment by:

@@

\paragraph{Synapses.}
\texttt{xolotl} provides some features for generating complex models. Synapses can be added with the \texttt{connect} function. At minimum synapses possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses. All specifiable properties of synapses are shown in Supplementary Table 2. To create axons or transport chains, the \texttt{slice} function splits a compartment into $n$ discrete segments and adds these compartments to the network connected by electrical synapses.

@@

All conductances contain fields for maximal conductance and reversal potential. Conductances with activation and inactivation variables include them as $m$ and $h$ respectively. Gating functions and their respective time constants are contained within the conductance header file. \texttt{xolotl} comes packaged with conductances from several dozen papers (Supplementary Table 3).

@@

\texttt{xolotl} contains template header files for producing custom conductances. The template contains instructions on how to design novel conductances with arbitrary specifications.

@@

Models are simulated in \texttt{xolotl} with the \texttt{integrate} function which outputs as time series the membrane potentials, intracellular calcium concentrations, controller states, intrinsic currents, and synaptic currents. The \texttt{integrate} function also accepts an argument which specifies injected current or clamped voltage.

@@

\paragraph{'Closed-loop' vs. 'open-loop.'}
Simulations can be run in 'closed-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or 'open-loop' mode which begins simulation with the current network state.

@@

\texttt{xolotl} comes packaged with a graphical user interface for visualizing parameter changes in real-time. The \texttt{manipulate} function opens the GUI, which displays a figure plotting the membrane potential and intracellular calcium concentration of all compartments as time series, and a dialog box with customizable sliders for all parameters of the model, much like the "Manipulate" function in Wolfram "Mathematica". Moving the sliders integrates the model in `open-loop` mode with the new parameters. The parameters available in the sliders can be customized by passing a cell array to manipulate. For example, to only see sliders for maximal conductances of the "HH" compartment, call "x.manipulate(x.find('HH*gbar'))". Closing the GUI saves the network state of the model to the "xolotl" object. This is particularly helpful for rapid prototyping of models.

@@

\paragraph{Limited to conductance-based models.} "xolotl" has been developed specifically for conductance-based models. It does not currently support current-based models.

\paragraph{Limited numerical integration strategies.} While the exponential Euler method performs well in neuronal models \autocite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}, it may be desirable to use other methods under certain conditions. "xolotl" does not currently support other integration schemes for its build-in conductances, nor does the software support error-sensitive variable step-sizes.

\paragraph{Inefficient tools for handling large networks.} While "xolotl" can integrate large networks ($>1000$ compartments), the tools used to index, label, hash, and search large networks become prohibitively slow. "xolotl" uses string-based comprehension for labeling compartments, which is suited to descriptively-named compartments, but computationally slow for searching and indexing.

@@

Using \texttt{xolotl} in \texttt{MATLAB}, users create a \texttt{xolotl} object and populate it with "cpplab" objects which describe compartments, conductances, synapses, and controllers. The model is integrated with the \texttt{integrate} function where the membrane potential, intracellular calcium concentration, controller states, intrinsic currents, and synaptic currents can be outputs.

\texttt{xolotl} comes packaged with a library of pre-existing conductance and synapse objects which greatly simplify the task of constructing model neurons. These objects can be referenced by name and added directly to a compartment. Novel conductance dynamics can be easily written by modifying a template header file contained in the \texttt{xolotl} distribution, or designed entirely from scratch.

@@

The seminal Hodgkin-Huxley model of action potentials in the squid giant axon \autocite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952} contains a fast inactivating sodium conductance ("NaV"), a non-inactivating delayed rectifier ("Kd"), and a passive leak current (Figure \ref{fig:figurehh}A). A compartment, \texttt{HH}, with membrane capacitance ("Cm") and surface area ("A") can be specified by Figure \ref{fig:figurehh}B. Network properties can be set during construction or afterwards using dot-notation in \texttt{MATLAB} (e.g. \texttt{x.HH.Cm}). Figure \ref{fig:figurehh}C shows the \texttt{MATLAB} command prompt after invoking the \texttt{xolotl} object \texttt{x}, displaying the hierarchical structure inherent in conductance-based treatments of neurodynamics.

This model was constructed using conductances from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion \autocite{turrigianoSelectiveRegulationCurrent1995}. In the absence of applied positive current, the model is quiescent. When 0.2 nA is injected, the model tonically spikes (Figure \ref{fig:figurehh}D). The \texttt{integrate} function takes the applied current as an argument (e.g. "x.integrate(Iapp)"), so that the \texttt{xolotl} object is agnostic to integration-specific perturbations. The \texttt{plot} function generates voltage and intracellular calcium traces, where the voltage trace is colored by the dominant current. If the membrane potential is increasing, the strongest instantaneous inward current colors the trace. Conversely, if the membrane potential is decreasing, the strongest outward current colors the trace instead. Figure \ref{fig:figurehh}F-I display the results of the \texttt{show} function. Activation and inactivation steady-states and the voltage-dependent time constants of these gating variables describe the conductance dynamics in absence of other channel types.

@@

\texttt{xolotl} can recapitulate the results of voltage clamp experiments \autocite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}. Figure Figure \ref{fig:figureclamp} displays steps in the procedure to clamp the membrane potential of a cell with delayed rectifier potassium conductance. During an \textit{in-vitro} experiment, confounding currents would be pharmacologically-blocked and two-electrode voltage clamp used to record tail currents at fixed membrane potential \autocite{connorInwardDelayedOutward1971, connorVoltageClampStudies1971}.

@@

Network models in \texttt{xolotl} consist of compartment objects connected by synapses. Synapses are stored in a vector array as a field of the \texttt{xolotl} object in \texttt{MATLAB}. Presynaptic and postsynaptic labels indicate the connectivity of the synapse. Figure Figure \ref{fig:figurenetwork} implements a model of the triphasic pyloric rhythm in the stomatogastric ganglion of crustaceans. The pyloric model contains three compartments and seven synapses (Figure \ref{fig:figurenetwork}A). This structure is reciprocated in the hierarchy of the \texttt{xolotl} object, where conductances are contained within compartments (Figure \ref{fig:figurenetwork}B).

@@

\texttt{xolotl} can implement homeostatic tuning rules as integral control. The controller computes an error signal (typically a function of intracellular calcium concentration), and adjusts the conductance or synapse it controls accordingly \autocite{olearyCorrelationsIonChannel2013}. In "xolotl", integral controllers are "cpplab" objects added to the conductance or synapse they regulate.

@@

\section*{Acknowledgments}
The authors would like to thank Mara CP Rue and Hillary Rodgers for beta-testing the \texttt{xolotl} software. Janis Li helped to prepare some conductance header files.

\section*{Supplemental Data}
Tables including all conductances packaged with \texttt{xolotl} should be put in the supplementary material.

@@

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_HH}
	\caption{\texttt{xolotl} can quickly set up and simulate conductance-based models. (A) Cartoon of a Hodgkin-Huxley single-compartment neuron model with fast sodium, delayed rectifier, and leak currents. (B) Code snippet in \texttt{MATLAB} used to implement D, F-I. (C) \texttt{xolotl} schematic displayed in the \texttt{MATLAB} command prompt. (D) Simulated voltage trace of a Hodgkin-Huxley model with three conductances and 0.2 nA of injected current. Colors indicate the dominant current (gold is fast sodium, blue is delayed rectifier, red is leak). (E) Firing rate-input relation displaying firing rate as a function of injected current current. (F-G) Steady-state gating functions for activation (m) and inactivation (h) gating variables. (H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables. Variables not plotted are unity for all voltage.}
	\label{fig:figurehh}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
	\caption{Simulating a network of conductance-based model neurons. (A) Diagram of a network model of the pyloric rhythm in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004). (B) Each neuron is modeled as a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances. (C) \textit{xolotl} implements conductances as fields of compartments and synapses as connections between compartments. (D-F) Simulated voltage trace of a model network for the three compartments. (G) Time series of synaptic currents in the simulated network can be obtained from the integration.}
	\label{fig:figurenetwork}
\end{figure}

>>>>

	"xolotl" is a free and open-source neuronal simulator written in \texttt{C++} with \texttt{MATLAB} wrappers. Biophysically-detailed models of networks can be designed efficiently using an intuitive language tightly coupled to the object-based architecture of the underlying \texttt{C++} code. Models can be specified by adding conductances to compartment objects. The structure is modular, serialized, and searchable, permitting high-level programmatic control over nearly all features of the models. \texttt{C++} templates are provided for developing new conductances, compartments, and integration schemata. It also includes a customizable graphical user interface (GUI) for rapid prototyping and hand-tuning conductances in real-time. The modular structure and accessibility to all parameters, variables, and dynamics of the model network in \texttt{MATLAB} facilitate rapid construction and assessment of model networks. "xolotl" is freely available at \url{https://github.com/marderlab/xolotl}. This tool provides straightforward implementation and fast simulation of neuronal models while permitting full control over every aspect of the network and integration.

@@

"xolotl" (\url{https://github.com/marderlab/xolotl}) is a fast single-compartment and multi-compartment simulator in \texttt{C++} with \texttt{MATLAB} wrappers. Written with an emphasis on ease-of-use, "xolotl" can simulate single-compartment conductance-based models, networks of these, and detailed multi-compartment models. "xolotl" exploits a novel automatic type system, \texttt{cpplab}, which binds \texttt{MATLAB} code to \texttt{C++} header files, creating objects and classes \textit{ad libitum} in \texttt{MATLAB} which reflect the underlying object-oriented code. "xolotl" implements \texttt{cpplab} to represent the nested structure of conductance-based models, and exploits the computational efficiency of the low-level programming language to quickly integrate models. For this reason, models can be implemented entirely in \texttt{MATLAB} with a few lines of code.

Models are specified in \texttt{MATLAB} by a nested structure. The "xolotl" object contains compartments which themselves contain conductances. Synapses belong to the "xolotl" object and connect compartments together. The high-level specification supports arbitrarily large network and multi-compartment morphologies.

The software has been implemented in \texttt{MATLAB} due to its ease-of-use and popularity among neuroscientists. \texttt{cpplab} provides a powerful backend for specifying and integrating models without relying on the significantly slower and limiting \texttt{MATLAB codegen}. While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.   Minimal experience with \texttt{MATLAB} is required to use "xolotl", and all equations and integration methods are provided transparently to the end user. No string parsing of equations is required \autocite{sherfeyDynaSimMATLABToolbox2018, stimbergBrianSecondComing2013, stimbergEquationorientedSpecificationNeural2014}.

"xolotl" comes packaged with visualization functions and a graphical user interface (GUI) for real-time manipulation of model parameters. Plotting of voltage, intracellular calcium, conductance gating functions, and time constants is provided by built-in "xolotl" methods. The GUI permits real-time tuning of any network parameters using numerical sliders in a graphical interface which displays the resultant membrane potential and intracellular calcium traces. The ease-of-use of these tools lends them to pedagogical applications and rapid exploration of toy models. This tool aims to simplify the investigation of dynamics of complex neural network models, facilitate collaborative modeling, and complement other tools being developed in the neuroinformatics community.

@@

\texttt{MATLAB} provides a high level programmatic and graphical interface for implementing, manipulating, and visualizing models without sacrificing the enhancements of the underlying \texttt{C++} code.

@@

Models are specified by adding compartments and synapses to the "xolotl" object. Conductances are added to compartments and controllers can be added to conductances. This modular structure recapitulates the biophysics of the Hodgkin-Huxley formalism and obviates the need to explicitly write out equations, which in "xolotl" are contained within the conductance header files.

@@

"xolotl" relies on \texttt{cpplab} constructions, which allow the user to exploit the efficiency of low-level \texttt{C++} code. \texttt{MATLAB} treats \texttt{cpplab} objects as fully-typed variables allowing for symbolic manipulation using only the high-level programming language and graphical interfaces. "xolotl" is fast because all time-intensive code is written in \texttt{C++}. While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming. \texttt{C++} provides speed improvements beyond the benefits of translating \texttt{MATLAB} features into low-level code. For this reason, \texttt{cpplab} has been designed to provide an interface for constructing, transpiling, and compiling \texttt{C++} code to be called from within \texttt{MATLAB}. "xolotl" simulations are run entirely from \texttt{C++} executables.

\paragraph{Automatic and efficient compiling.}
"xolotl" automatically handles transpiling and compiling "MATLAB" code into "C++". The MD5 algorithm hashes the network to compile a new binary and "MEX" bridge file only if needed and to confirm that the correct binary fetched during execution.

@@

\paragraph{Adding features.}
The "add" function creates a compartment, conductance, or controller and affixes it as a field in the "xolotl" network structure. This function generates a "MATLAB struct" that faithfully represents the underlying "C++" code. Compartments add to the "xolotl" object and conductances add to compartments. Specific properties can be specified using key-value pair arguments (e.g. Figure \ref{fig:figurehh}A).

\paragraph{Finding features.}
\texttt{cpplab} comes with several features which simplify the handling of complexly-nested models. The "find" function acquires a cell array of all properties of the network which satisfy a search condition. For example, one can find all paths to maximal conductances within the "'HH'" compartment by:

@@

\paragraph{Synapses.}
"xolotl" provides some features for generating complex models. Synapses can be added with the \texttt{connect} function. At minimum synapses possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses. All specifiable properties of synapses are shown in Supplementary Table 2. To create axons or transport chains, the \texttt{slice} function splits a compartment into $n$ discrete segments and adds these compartments to the network connected by electrical synapses.

@@

All conductances contain fields for maximal conductance and reversal potential. Conductances with activation and inactivation variables include them as $m$ and $h$ respectively. Gating functions and their respective time constants are contained within the conductance header file. "xolotl" comes packaged with conductances from several dozen papers (Supplementary Table 3).

@@

"xolotl" contains template header files for producing custom conductances. The template contains instructions on how to design novel conductances with arbitrary specifications.

@@

Models are simulated in "xolotl" with the \texttt{integrate} function which outputs as time series the membrane potentials, intracellular calcium concentrations, controller states, intrinsic currents, and synaptic currents. The \texttt{integrate} function also accepts an argument which specifies injected current or clamped voltage.

@@

\paragraph{`Closed-loop' vs. `open-loop.'}
Simulations can be run in `closed-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or `open-loop' mode which begins simulation with the current network state.

@@

"xolotl" comes packaged with a graphical user interface for visualizing parameter changes in real-time. The \texttt{manipulate} function opens the GUI, which displays a figure plotting the membrane potential and intracellular calcium concentration of all compartments as time series, and a dialog box with customizable sliders for all parameters of the model, much like the "Manipulate" function in Wolfram "Mathematica". Moving the sliders integrates the model in `open-loop` mode with the new parameters. The parameters available in the sliders can be customized by passing a cell array to manipulate. For example, to only see sliders for maximal conductances of the "HH" compartment, call "x.manipulate('HH*gbar')". Closing the GUI saves the network state of the model to the "xolotl" object. This is particularly helpful for rapid prototyping of models.

@@

\paragraph{Limited to conductance-based models.} "xolotl" has been developed specifically for conductance-based models. It does not currently support rate- or current-based models.

\paragraph{Limited numerical integration strategies.} While the exponential Euler method performs well in neuronal models \autocite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}, it may be desirable to use other methods under certain conditions. "xolotl" does not currently support other integration schemes for its built-in conductances, nor does the software support error-sensitive variable step-sizes `out-of-the-box.'

\paragraph{Inefficient tools for handling large networks.} While "xolotl" can integrate large networks ($>1000$ compartments), "xolotl" uses string-based comprehension for labeling compartments which is suited to descriptive naming, but prohibits vector operations over compartments.

@@

In "MATLAB", users create a "xolotl" object and populate it with "cpplab"-generated objects which describe compartments, conductances, synapses, and controllers. The model is integrated with the "integrate" function where the membrane potential, intracellular calcium concentration, controller states, intrinsic currents, and synaptic currents can be outputs.

"xolotl" comes packaged with a library of pre-existing conductance and synapse objects which greatly simplify the task of constructing model neurons. These objects can be referenced by name and added directly to a compartment. Novel conductance dynamics can be easily written by modifying a template header file contained in the "xolotl" distribution, or designed entirely from scratch.

@@

The seminal Hodgkin-Huxley model of action potentials in the squid giant axon \autocite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952} contains a fast inactivating sodium conductance ("NaV"), a non-inactivating delayed rectifier ("Kd"), and a passive leak current (Figure \ref{fig:figurehh}A). A compartment, \texttt{HH}, with membrane capacitance ("Cm") and surface area ("A") can be specified by Figure \ref{fig:figurehh}B. Network properties can be set during construction or afterwards using dot-notation in \texttt{MATLAB} (e.g. \texttt{x.HH.Cm}). Figure \ref{fig:figurehh}C shows the \texttt{MATLAB} command prompt after invoking the "xolotl" object \texttt{x}, displaying the hierarchical structure inherent in conductance-based treatments of neurodynamics.

This model was constructed using conductances from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion \autocite{turrigianoSelectiveRegulationCurrent1995}. In the absence of applied positive current, the model is quiescent. When 0.2 nA is injected, the model tonically spikes (Figure \ref{fig:figurehh}D). The \texttt{integrate} function takes the applied current as an argument (e.g. "x.integrate(Iapp)"), so that the "xolotl" object is agnostic to integration-specific perturbations. The \texttt{plot} function generates voltage and intracellular calcium traces, where the voltage trace is colored by the dominant current. If the membrane potential is increasing, the strongest instantaneous inward current colors the trace. Conversely, if the membrane potential is decreasing, the strongest outward current colors the trace instead. Figure \ref{fig:figurehh}F-I display the results of the \texttt{show} function. Activation and inactivation steady-states and the voltage-dependent time constants of these gating variables describe the conductance dynamics in absence of other channel types.

@@

"xolotl" can recapitulate the results of voltage clamp experiments \autocite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}. Figure Figure \ref{fig:figureclamp} displays steps in the procedure to clamp the membrane potential of a cell with delayed rectifier potassium conductance. During an \textit{in-vitro} experiment, confounding currents would be pharmacologically-blocked and two-electrode voltage clamp used to record tail currents at fixed membrane potential \autocite{connorInwardDelayedOutward1971, connorVoltageClampStudies1971}.

@@

Network models in "xolotl" consist of compartment objects connected by synapses. Synapses are stored in a vector array as a field of the "xolotl" object in \texttt{MATLAB}. Presynaptic and postsynaptic labels indicate the connectivity of the synapse. Figure Figure \ref{fig:figurenetwork} implements a model of the triphasic pyloric rhythm in the stomatogastric ganglion of crustaceans. The pyloric model contains three compartments and seven synapses (Figure \ref{fig:figurenetwork}A). This structure is reciprocated in the hierarchy of the "xolotl" object, where conductances are contained within compartments (Figure \ref{fig:figurenetwork}B).

@@

"xolotl" can implement homeostatic tuning rules as integral control. The controller computes an error signal (typically a function of intracellular calcium concentration), and adjusts the conductance or synapse it controls accordingly \autocite{olearyCorrelationsIonChannel2013}. In "xolotl", integral controllers are "cpplab" objects added to the conductance or synapse they regulate.

@@

\section*{Acknowledgments}
The authors would like to thank Mara CP Rue and Hillary Rodgers for beta-testing the "xolotl" software. Janis Li helped to prepare some conductance header files.

\section*{Supplemental Data}
Tables including all conductances packaged with "xolotl" should be put in the supplementary material.

@@

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_HH}
	\caption{\texttt{xolotl} can quickly set up and simulate conductance-based models. (A) Cartoon of a Hodgkin-Huxley single-compartment neuron model with fast sodium, delayed rectifier, and leak currents. (B) Code snippet in \texttt{MATLAB} used to implement D, F-I. (C) \texttt{xolotl} schematic displayed in the \texttt{MATLAB} command prompt. (D) Simulated voltage trace of a Hodgkin-Huxley model with three conductances and 0.2 nA of injected current. Colors indicate the dominant current (gold is fast sodium, blue is delayed rectifier, red is leak). (E) Firing rate-input relation displaying firing rate as a function of injected current current. (F-G) Steady-state gating functions for activation (m) and inactivation (h) gating variables. (H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables.}
	\label{fig:figurehh}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
	\caption{Simulating a network of conductance-based model neurons. (A) Diagram of a network model of the pyloric rhythm in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004). (B) Each neuron is modeled as a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances. (C) \texttt{xolotl} implements conductances as fields of compartments and synapses as connections between compartments. (D-F) Simulated voltage trace of a model network for the three compartments. (G) Time series of synaptic currents in the simulated network can be obtained from the integration.}
	\label{fig:figurenetwork}
\end{figure}
